---
title: "EDA and Time Series Forecasting of Avocado prices"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DS 5110 - INTRODUCTION TO DATA MANAGEMENT AND PROCESSING
## GROUP 3 - FINAL PROJECT CODE
## EDA and Time Series Forecasting of Avocado prices
## Authors: Amey Shankar Basangoudar, Priyesh Priyesh, Riddhi Narayan, Saachi Chandrashekhar
## Dataset: https://www.kaggle.com/neuromusic/avocado-prices


```{r}
#Importing libraries

library(tidyverse)
library(modelr)
library(repr)
library(grid)
library(dplyr)
library(cowplot) 
library(forecast)
library(ggplot2)
library(grid)
library(corrplot)
library(RColorBrewer)
library(Metrics)
library(caret)
```


```{r}
#Importing dataset
df <- read_csv("E:/SAACHI/PERSONAL/COLLEGES/NEU/2021/IDMP/PROJECT/Avocado/avocado.csv", na = ".")
head(df)
```


```{r}
#Tidying the data
avocado_tidied <- df %>% pivot_longer(cols = c(`4046`, `4225`, `4770`), names_to = "PLU type", values_to = "Volume of each PLU Sold")
avocado_tidied <- avocado_tidied %>% pivot_longer(cols = c(`Small Bags`, `Large Bags`, `XLarge Bags`), names_to = "Bag type", values_to = "Volume of each bag sold")
```


```{r}
#Distribution of all variables 
options(scipen = 999)
library(grid)
p1 <- avocado_tidied %>% 
  ggplot(aes(x = AveragePrice)) + 
  geom_histogram(aes(y = ..density..), color = 'black', fill = 'white') +
  geom_density(alpha=.2, fill="#FF6666") + theme_dark()

p2 <- avocado_tidied %>% 
  ggplot(aes(x = `Total Volume`)) + 
  geom_boxplot(fill="#FF6666") + 
  scale_x_log10() +
  theme_dark() + 
  coord_flip()

p3 <- avocado_tidied %>% 
  ggplot(aes(x = `Volume of each PLU Sold`)) + 
  geom_boxplot(fill="#FF6666") + 
  facet_wrap(~`PLU type`) +
  scale_x_log10() +
  theme_dark() + 
  coord_flip()

p4 <- avocado_tidied %>% 
  ggplot(aes(x = `Total Bags`)) + 
  geom_boxplot(fill="yellow") + 
  scale_x_log10() +
  theme_dark() + 
  coord_flip()

p5 <- avocado_tidied %>% 
  ggplot(aes(x = `Volume of each bag sold`)) + 
  geom_boxplot(fill="yellow") + 
  facet_wrap(~`Bag type`) +
  scale_x_log10() +
  theme_dark() + 
  coord_flip()



vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)

grid.newpage()
pushViewport(viewport(layout = grid.layout(8, 3))) # 5 rows, 3 columns
print(p1, vp = vplayout(1:4, 1:2))  # the big plot covers rows 1:2 and cols 1:3
print(p2, vp = vplayout(5:6, 1))
print(p3, vp = vplayout(5:6, 2:3))
print(p4, vp = vplayout(7:8, 1))
print(p5, vp = vplayout(7:8, 2:3))
```

## EXPLORATORY DATA ANALYSIS

```{r}
#Price per type
ggplot(df, aes(df$type, df$AveragePrice, color = df$type))+
    geom_boxplot()+
    labs(x = "Type", y ="Average Price", title = "Average price of avocado by avocado type")
```

```{r}
# 1. Distribution of average price for each region 
options(repr.plot.width=15, repr.plot.height=8)
avocado_tidied %>% 
  ggplot(aes(x = region, y = AveragePrice)) + 
  geom_boxplot(aes(fill = region), show.legend = FALSE) + 
  theme(axis.text.x = element_text(angle = 90)) 
```


```{r}
# 2. Average Price Vs Total volume sold 
avocado_tidied %>% 
  ggplot(aes(x = `Total Volume`, y = AveragePrice)) +
  geom_point(aes(color = type), alpha = 0.05) +
  geom_smooth(aes(color = type)) + 
  scale_x_log10()
```


```{r}
# 3. Distribution of AveragePrice avocados in each region for each type 
options(repr.plot.width=20, repr.plot.height=8)
avocado_tidied %>% 
  ggplot(aes(x = region, y = AveragePrice)) + 
  geom_boxplot(aes(fill = `type`)) + 
  theme(axis.text.x = element_text(angle = 90))
```


```{r}
# 4. Distribution of  avocados in each region for each type 

options(repr.plot.width=20, repr.plot.height=8)
avocado_tidied %>% 
  ggplot( aes(x = region, y = `Total Volume`)) + 
  geom_boxplot(aes(fill = `type`)) + 
  scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90))
```


```{r}
# 5. AveragePrice Vs Total volume sold in each year
avocado_tidied %>% 
  ggplot(aes(x = `Total Volume`, y = AveragePrice)) +
  geom_hex(aes(color = as.factor(year)), alpha = 0.03, show.legend = FALSE) +
  geom_smooth(aes(color = as.factor(year))) +
  facet_wrap(~year) +
  scale_x_log10()
```


```{r}
# 6. Price vs date (for each type of avocado) 

ggplot(data = df, aes(x = Date, y = AveragePrice, color=type)) +
     geom_line() + ggtitle('Distribution of price vs. date') +
     facet_wrap(~ type) + theme_minimal() + theme(legend.position="right")

ggplot(data = df, aes(x = Date, y = `Total Volume`, color=type)) +
     geom_line() + ggtitle('Distribution of price vs. date') +
     facet_wrap(~ type) + theme_minimal() + theme(legend.position="right")
```


```{r}
# 7. Change of price with repect to date
df1 <- df %>% select(Date,AveragePrice)  %>% group_by(Date)  %>% summarise(AveragePrice = mean(AveragePrice), n = n())%>% arrange(Date) 
head(df1)

ggplot() + geom_line(data=df1, aes(x =Date, y=(AveragePrice)
                                         ))  +
  labs(x="Date", y="price",
       title="Change of price with repect to date") + 
  theme(axis.text.x = element_text(angle = 90))
```


```{r}
# 8. Count per region
avoc <- df %>%
  rename(volume = `Total Volume`, total_bag = `Total Bags`)
 
avoc %>%
  group_by(region) %>%
  summarise(volume = sum(volume)) %>%
  ggplot(aes(reorder(region, volume), volume)) +
  geom_col( width = 0.5) +
 coord_flip()+
  labs(title = "Count of Avocados by region",
       x = "", y = "Total volume")
```


```{r}
# 9. Price per region
df %>%
  group_by(region) %>%
  summarise(price = mean(AveragePrice)) %>%
  arrange(price) %>%
  ggplot(aes(reorder(region, price), price)) +
  geom_point() +
  coord_flip() +
  labs(title = "Avg avocado price by region",
        y = "Cost", x = "Region") 
```


```{r}
# 10. Avg Price and Sales in Boston

Bos <- avoc %>% filter(df$region == "Boston")

ggplot(data = Bos, aes(x = Date, y = AveragePrice, color=type)) +
     geom_line() + ggtitle('BOSTON - Distribution of price vs. date') +
      theme_minimal() + theme(legend.position="right")

ggplot(data = Bos, aes(x = Date, y = volume, color=type)) +
     geom_line() + ggtitle('BOSTON - Distribution of volume vs. date') +
      theme_minimal() + theme(legend.position="right")
```


```{r}
# 11. Bags sold
mymonths <- c("Jan","Feb","Mar",
              "Apr","May","Jun",
              "Jul","Aug","Sep",
              "Oct","Nov","Dec")
#Distribution of average sales of different sizes per month
df_date <- df%>% mutate(year = lubridate::year(Date), 
                month = lubridate::month(Date), 
                day = lubridate::day(Date))

df_date$Month_Name <- mymonths[ df_date$month ]

df_new <- df_date%>%
rename(small = `Small Bags`, large = `Large Bags`,xlarge = `XLarge Bags`)
head(df_new)

bag_data<- df_new%>%
group_by(Date,year,Month_Name,month) %>%
summarize(s_bag = mean(small),l_bag = mean(large),
xl_bag = mean (xlarge),avg_price = mean(AveragePrice ))%>%
arrange(Date)

plot_bag_data <- bag_data %>% gather(key = "bagsize", value="Total_no_of_bags", c(5:7))

plot_bag_data$month_abb <- sapply(plot_bag_data$month, function(x) month.abb[as.numeric(x)])
plot_bag_data$month_abb = factor(plot_bag_data$month_abb, levels = month.abb)

options(repr.plot.width=15, repr.plot.height=8)
plot_bag_data$bagsize  <- factor(plot_bag_data$bagsize, levels=c("xl_bag", "l_bag", "s_bag"))

ggplot(plot_bag_data, aes(x= month_abb, y=Total_no_of_bags, color = bagsize))+
    geom_bar(aes(x= month_abb, y=Total_no_of_bags, fill = bagsize), stat= 'identity',color="white")+
    scale_fill_manual(values = c("purple", "orange", "red")) +
    labs(y= "Total Bag Sales" ,x="Month",title = "Sales in 2015-2018")+
    theme(legend.position = "bottom",axis.text.y = element_text( angle=45),
    axis.text.x= element_text( angle=90),
    legend.direction = "horizontal")  +
    facet_grid(cols=vars(year))
```

```{r}
# 12. Choropleth
avocado_df_new <- df
us_cities <- read_csv("citystate.csv", na = ".")

 MainStates <- map_data("state")

 MainStates1 <- MainStates %>% 
  rename(
    State = region
    )
 df1 <- MainStates1 %>% select((State))
 distinct(df1)
 avocado_df2 <- avocado_df_new   %>% select(region,AveragePrice,Date)  %>% mutate(region = tolower(region))  %>% group_by(region,Date) %>% summarise(AveragePrice = mean(AveragePrice), n = n())
abc<-avocado_df2 %>% select(region) %>% distinct()  %>% print()
#view(abc)
 us_cities1 <- us_cities  %>% mutate(region = tolower(city),State = tolower(state) )
us_cities1<- distinct(us_cities1) %>% arrange(region) 
 avocado_df_state <- inner_join(us_cities1,avocado_df2, sort = Date,by="region")

choro <- inner_join(MainStates1,avocado_df_state, sort = FALSE, by = "State") 
 
 choro <- choro %>% mutate(year=format(Date,format="%Y"))

 


ggplot(MainStates,aes(long, lat, map_id = region)) + geom_map(map = MainStates,color = "black",size = 0.3,fill="grey") + geom_polygon( data=choro , aes(x=long, y=lat, group=State,
                  fill=AveragePrice), color="black" ) + facet_wrap(~year) +coord_map()
```


## MODELLING

## LINEAR REGRESSION
```{r}
#AVERAGE PRICE VS VOLUME

ggplot(avoc, aes(x=log10(volume), y=AveragePrice)) +
  geom_point(size = 1) + geom_smooth(method="lm") +
  labs(title = "Avg price vs vol", 
       x = "Volume" ,
       y = "Avg Price") +
   theme_minimal()

ggplot(avoc, aes(x=log10(volume), y=log10(AveragePrice))) +
  geom_point(size = 1) + geom_smooth(method="lm") +
  labs(title = "Avg price vs vol", 
       x = "Volume" ,
       y = "Avg Price") +
   theme_minimal()


#AVERAGE PRICE VS TOTAL BAGS

ggplot(avoc, aes(x=log10(total_bag), y=AveragePrice)) +
  geom_point(size = 1) + geom_smooth(method="lm") +
  labs(title = "Avg price vs Total bags", 
       x = "Total bags" ,
       y = "Avg Price") +
   theme_minimal()

ggplot(avoc, aes(x=log10(total_bag), y=log10(AveragePrice))) +
  geom_point(size = 1) + geom_smooth(method="lm") +
  labs(title = "Avg price vs Total bags", 
       x = "Total bags" ,
       y = "Avg Price") +
   theme_minimal()


#AVERAGE PRICE VS 4046 type

ggplot(avoc, aes(x=log10(`4046`), y=AveragePrice)) +
  geom_point(size = 1) + geom_smooth(method="lm") +
  labs(title = "Avg price vs 4046 type", 
       x = "4046 type" ,
       y = "Avg Price") +
   theme_minimal()

ggplot(avoc, aes(x=log10(`4046`), y=log10(AveragePrice))) +
  geom_point(size = 1) + geom_smooth(method="lm") +
  labs(title = "Avg price vs 4046 type", 
       x = "4046 type" ,
       y = "Avg Price") +
   theme_minimal()
  

#AVERAGE PRICE VS 4225 type

ggplot(avoc, aes(x=log10(`4225`), y=AveragePrice)) +
  geom_point(size = 1) + geom_smooth(method="lm") +
  labs(title = "Avg price vs 4225 type", 
       x = "4225 type" ,
       y = "Avg Price") +
   theme_minimal()

ggplot(avoc, aes(x=log10(`4225`), y=log10(AveragePrice))) +
  geom_point(size = 1) + geom_smooth(method="lm") +
  labs(title = "Avg price vs 4225 type", 
       x = "4225 type" ,
       y = "Avg Price") +
   theme_minimal()


#AVERAGE PRICE VS 4770 type

ggplot(avoc, aes(x=log10(`4770`), y=AveragePrice)) +
  geom_point(size = 1) + geom_smooth(method="lm") +
  labs(title = "Avg price vs 4770 type", 
       x = "4770 type" ,
       y = "Avg Price") +
   theme_minimal()

ggplot(avoc, aes(x=log10(`4770`), y=log10(AveragePrice))) +
  geom_point(size = 1) + geom_smooth(method="lm") +
  labs(title = "Avg price vs 4770 type", 
       x = "4770 type" ,
       y = "Avg Price") +
   theme_minimal()

#AVERAGE PRICE VS TYPE

ggplot(avoc, aes(x=type, y=AveragePrice, fill=type)) +
  geom_boxplot() +
  scale_fill_brewer(palette="Set1", direction=-1) +
  labs(y="Average Price",
       title="Average price vs Type") +
  theme_minimal()

#AVERAGE PRICE VS REGION

ggplot(avoc, aes(x=region, y=AveragePrice, fill=region)) +
  geom_boxplot() +
  labs(y="Average Price",
       title="Average price vs Region") +
  theme_minimal()
```


```{r}
# FITS

fit_1 <- lm(log10(AveragePrice) ~ log10(volume), data=avoc)
fit_2 <- lm(log10(AveragePrice) ~ type, data=avoc)
fit_3 <- lm(log10(AveragePrice) ~ region, data=avoc)
fit_4 <- lm(log10(AveragePrice) ~ log10(total_bag+1), data=avoc)
fit_5 <- lm(log10(AveragePrice) ~ log10(`4046` + 1), data=avoc)
fit_6 <- lm(log10(AveragePrice) ~ log10(`4225` + 1), data=avoc)
fit_7 <- lm(log10(AveragePrice) ~ log10(`4770`+ 1), data=avoc)

summary(fit_1)
summary(fit_2)
summary(fit_3)
summary(fit_4)
summary(fit_5)
summary(fit_6)
summary(fit_7)
```


```{r}
#RESIDUAL

avoc %>%
  add_residuals(fit_1, "resid") %>%
  ggplot(aes(x=log10(volume))) +
  geom_point(aes(y=resid), alpha=0.5) +
  labs(x="Log volume", y="Residuals") +
  theme_minimal()

avoc %>%
  add_residuals(fit_2, "resid") %>%
  ggplot(aes(x=type)) +
  geom_boxplot() +
  labs(x="Type", y="Residuals") +
  theme_minimal()

avoc %>%
  add_residuals(fit_3, "resid") %>%
  ggplot(aes(x=region)) +
  geom_boxplot() +
  labs(x="Region", y="Residuals") +
  theme_minimal()

avoc %>%
  add_residuals(fit_4, "resid") %>%
  ggplot(aes(x=log10(total_bag))) +
  geom_point(aes(y=resid), alpha=0.5) +
  labs(x="Log total bag", y="Residuals") +
  theme_minimal()

avoc %>%
  add_residuals(fit_5, "resid") %>%
  ggplot(aes(x=log10(`4046`))) +
  geom_point(aes(y=resid), alpha=0.5) +
  labs(x="Log 4046", y="Residuals") +
  theme_minimal()

avoc %>%
  add_residuals(fit_6, "resid") %>%
  ggplot(aes(x=log10(`4225`))) +
  geom_point(aes(y=resid), alpha=0.5) +
  labs(x="Log 4225", y="Residuals") +
  theme_minimal()

avoc %>%
  add_residuals(fit_7, "resid") %>%
  ggplot(aes(x=log10(`4770`))) +
  geom_point(aes(y=resid), alpha=0.5) +
  labs(x="Log 4770", y="Residuals") +
  theme_minimal()
```


```{r}
# QQ PLOT

avoc %>%
  add_residuals(fit_1, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() + 
  theme_minimal()

avoc %>%
  add_residuals(fit_2, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal()

avoc %>%
  add_residuals(fit_3, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal()

avoc %>%
  add_residuals(fit_4, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal()

avoc %>%
  add_residuals(fit_5, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal()

avoc %>%
  add_residuals(fit_6, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal()

avoc %>%
  add_residuals(fit_7, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal()
```


```{r}
#AIC

first_model <- lm(log10(AveragePrice) ~ region, data=avoc)
fit_all <- lm(log10(AveragePrice) ~ log10(volume) + log10(total_bag+1) + type + region + log10(`4046`+1) + log10(`4225`+1) + log10(`4770`+1) , data = avoc)
step(first_model, direction = 'forward', scope = formula(fit_all))
```


```{r}
# AIC

first_model <- lm(log10(AveragePrice) ~ region, data=avoc)
fit_all <- lm(log10(AveragePrice) ~ log10(volume) + log10(total_bag+1) + type + region + log10(`4046`+1) + log10(`4225`+1) + log10(`4770`+1) , data = avoc)
step(first_model, direction = 'forward', scope = formula(fit_all))
```


```{r}
#RMSE


set.seed(2)
av_part <- resample_partition(avoc,
                                    p=c(train=0.5,
                                        valid=0.25,
                                        test=0.25))

av_part
```


```{r}
step1 <- function(response, predictors, candidates, partition)
{
  rhs <- paste0(paste0(predictors, collapse="+"), "+", candidates)
  formulas <- lapply(paste0(response, "~", rhs), as.formula)
  rmses <- sapply(formulas,
                  function(fm) rmse(lm(fm, data=partition$train),
                                    data=partition$valid))
  names(rmses) <- candidates
  attr(rmses, "best") <- rmses[which.min(rmses)]
  rmses
}

model <- NULL
```


```{r}
#Step 1 (no variables):

preds <- "1"
cands <- c("log10(volume)" , "log10(total_bag+1)" , "type" , "region" , 
"log10(`4046`+1)" , "log10(`4225`+1)" , "log10(`4770`+1)")
s1 <- step1("log10(AveragePrice)", preds, cands, av_part)

model <- c(model, attr(s1, "best"))
s1
```


```{r}
#Step 2:

preds <- "type"
cands <- c("log10(volume)" , "log10(total_bag+1)" ,  "region" , "log10(`4046`+1)" , "log10(`4225`+1)" , "log10(`4770`+1)")
s1 <- step1("log10(AveragePrice)", preds, cands, av_part)

model <- c(model, attr(s1, "best"))
s1
```


```{r}
#Step 3:

preds <- c("type", "region")
cands <- c("log10(volume)" , "log10(total_bag+1)" , "log10(`4046`+1)" , "log10(`4225`+1)" , "log10(`4770`+1)")
s1 <- step1("log10(AveragePrice)", preds, cands, av_part)

model <- c(model, attr(s1, "best"))
s1
```


```{r}
#Step 4: 

preds <- c("type", "region", "log10(volume)")
cands <- c( "log10(total_bag+1)" , "log10(`4046`+1)" , "log10(`4225`+1)" , "log10(`4770`+1)")
s1 <- step1("log10(AveragePrice)", preds, cands, av_part)

model <- c(model, attr(s1, "best"))
s1
```


```{r}
#Step 5:


preds <- c("type", "region", "log10(volume)", "log10(`4225`+1)")
cands <- c( "log10(total_bag+1)" , "log10(`4046`+1)" ,  "log10(`4770`+1)")
s1 <- step1("log10(AveragePrice)", preds, cands, av_part)

model <- c(model, attr(s1, "best"))
s1
```

```{r}
#Step 6: 


preds <- c("type", "region", "log10(volume)", "log10(`4225`+1)","log10(`4770`+1)")
cands <- c( "log10(total_bag+1)" , "log10(`4046`+1)")
s1 <- step1("log10(AveragePrice)", preds, cands, av_part)

model <- c(model, attr(s1, "best"))
s1
```


```{r}
#Step 7:

preds <- c("type", "region", "log10(volume)", "log10(`4225`+1)","log10(`4770`+1)", "log10(total_bag+1)")
cands <- c(  "log10(`4046`+1)")
s1 <- step1("log10(AveragePrice)", preds, cands, av_part)

model <- c(model, attr(s1, "best"))
s1
```


```{r}
step_model <- tibble(index=seq_along(model),
                     variable=factor(names(model), levels=unique(names(model))),
                     RMSE=model)

ggplot(step_model, aes(y=RMSE)) +
  geom_point(aes(x=variable)) + 
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  geom_line(aes(x=index)) +
  labs(title="Stepwise model selection") +
  theme_minimal()
```


```{r}
# LINEAR REGRESSION MODEL
set.seed(1)
row.number <- sample(1:nrow(avoc), 0.8*nrow(avoc))
train = avoc[row.number,]
test = avoc[-row.number,]
dim(train)
dim(test)
```


```{r}
Final_fit <- lm(formula = log10(AveragePrice) ~ type + region + log10(volume) + 
    log10(`4225` + 1) + log10(`4770` + 1) + log10(total_bag + 
    1) + log10(`4046` + 1), data = train)

summary(Final_fit)
plot(Final_fit)
```


```{r}
pred <- predict(Final_fit, data = test)

actuals_preds <- data.frame(cbind(actuals=test$AveragePrice, predicteds=pred))
linear_df <- data.frame(cbind(AveragePrice=test$AveragePrice, pred=pred))

correlation_accuracy <- cor(actuals_preds)
correlation_accuracy

head(actuals_preds)

rmse(actuals_preds$actuals,actuals_preds$predicteds)
```

## RANDOM FOREST
```{r}
#Random Forest

avoc = avoc %>% mutate(id = rownames(avoc))

set.seed(46848)
test_rows = sample(avoc$id, nrow(avoc)*0.2, replace = F) %>% as.numeric()

train_rf = avoc %>% filter(!id %in% test_rows)
test_rf = avoc %>% filter(id %in% test_rows) %>% select(-AveragePrice)

head(train_rf)
dim(train_rf)

head(test_rf)
dim(test_rf)
```


```{r}
rf = train( AveragePrice ~.,  data = train_rf %>% select(-id),
    method = "ranger",
    num.trees = 100,
    trControl = trainControl(method = "oob"),
    tuneGrid = expand.grid("mtry" = c(8, 10, 12, 14),
                          "splitrule" = "variance",
                          "min.node.size" = 1:10)
)
```


```{r}
rf_pred = predict(
    rf,
    newdata = test_rf
)

# The results
rf_df = data.frame(
    id = test_rf %>% select(id),
    AveragePrice = avoc %>% filter(id %in% test_rows) %>% select(AveragePrice),
    pred = rf_pred
)
```

```{r}
rmse = function(df){
    (df$AveragePrice - df$pred)^2 %>% mean() %>% sqrt()
}

rmse(rf_df)
rmse(linear_df)
```


```{r}
all_data <- data.frame(model_name = c("random forest", "linear regression"), 
rmse_vals = c(rmse(rf_df), rmse(linear_df)))

ggplot(all_data, aes(model_name, rmse_vals)) + 
geom_col(fill = c("darkgreen"), col = "red") + 
geom_text(aes(label = round(rmse_vals, 4)), vjust = 0)+ 
ggtitle("RMSE Values") + 
theme_light()
```






## TIME SERIES ANALYSIS
```{r}
library(fpp2)
library(tibbletime)
library(dplyr)

```


```{r}
# Let's declare our data as time series
conv <- df%>% select(Date, AveragePrice, type) %>% filter(type == "conventional")
org <- df %>% select(Date, AveragePrice, type) %>% filter(type == "organic")
```


```{r}

# conventional 
conventional <- as_tbl_time(conv, index=Date)
conventional<- as_period(conventional, '1 month')
conventional$type <- NULL

# Organic Avocados
organic <- as_tbl_time(org, index=Date)
organic <- as_period(organic, '1 month')
organic$type <- NULL
```


```{r}
conv_ts <- ts(conventional[,2], start=c(2015, 1),end=c(2018,12), frequency=12)
org_ts <- ts(organic[,2], start=c(2015, 1),end=c(2018,12), frequency=12)
differences_conv <- diff(conv_ts)

main_diff <- autoplot(differences_conv) + theme_minimal()
seasonality_diff <- ggseasonplot(differences_conv) + theme_minimal()

plot_grid(main_diff, seasonality_diff, nrow=2)
head(conventional)
```


```{r}
arima_model_cv <- auto.arima(conv_ts, d=1, D=1, stepwise=FALSE, approximation=FALSE, trace=TRUE)
arima_model_or <- auto.arima(org_ts, d=1, D=1, stepwise=FALSE, approximation=FALSE, trace=TRUE)


print(summary(arima_model_cv))
checkresiduals(arima_model_cv) + theme_minimal()

```


```{r}
print(summary(arima_model_or))
checkresiduals(arima_model_or) + theme_minimal()
```


```{r}
options(repr.plot.width=10, repr.plot.height=7)
install.packages('ggthemes')
library(ggthemes)

conv_forecast_sn <- autoplot(conv_ts) +
  autolayer(meanf(conv_ts, h=24),
    series="Mean", PI=FALSE) +
  autolayer(naive(conv_ts, h=24),
    series="Na誰ve", PI=FALSE) +
  autolayer(snaive(conv_ts, h=24),
    series="Seasonal na誰ve", PI=FALSE) +
  ggtitle("Conventional Avocado \n Seasonal Naive Method") +
  xlab("Date") + ylab("Price") + scale_color_manual(values=c("#FA5858", "#00BFFF", "#FF8000")) + 
  guides(colour=guide_legend(title="Forecast"))  + theme_economist() + 
theme(plot.title=element_text(hjust=0.5), plot.background=element_rect(fill="#F4F6F7"))

org_forecast_sn <- autoplot(org_ts) +
  autolayer(meanf(org_ts, h=24),
    series="Mean", PI=FALSE) +
  autolayer(naive(org_ts, h=24),
    series="Na誰ve", PI=FALSE) +
  autolayer(snaive(org_ts, h=24),
    series="Seasonal na誰ve", PI=FALSE) +
  ggtitle("Organic Avocado \n Seasonal Naive Method") +
  xlab("Date") + ylab("Price") + scale_color_manual(values=c("#FA5858", "#00BFFF", "#FF8000")) + 
  guides(colour=guide_legend(title="Forecast"))  + theme_economist() + 
theme(plot.title=element_text(hjust=0.5), plot.background=element_rect(fill="#DCFCE6"), legend.position="none")


plot_grid(conv_forecast_sn, org_forecast_sn, nrow=2)
#plot(conv_forecast_sn)
```


```{r}
forecast_cv <- forecast(arima_model_cv, h=24)
# Include means including the last 60 months in order to see closer the forecast.
autoplot(forecast_cv, include=60) + theme_minimal() + theme(plot.title=element_text(hjust=0.5), plot.background=element_rect(fill="#F4F6F7"),
      legend.position="bottom", legend.background = element_rect(fill="#FFF9F5",
                                  size=0.5, linetype="solid", 
                                  colour ="black")) + 
labs(title="Forecasting using ARIMA model \n Conventional Avocados", x="Date", y="Price")
```


```{r}
forecast_org <- forecast(arima_model_or, h=24)
# Include means including the last 60 months in order to see closer the forecast.
autoplot(forecast_org, include=60) + theme_minimal() + theme(plot.title=element_text(hjust=0.5), plot.background=element_rect(fill="#d0f0c0"),
      legend.position="bottom", legend.background = element_rect(fill="#fffacd",
                                  size=0.5, linetype="solid", 
                                  colour ="black")) + 
labs(title="Forecasting using ARIMA model \n Organic Avocados", x="Date", y="Price")

```


```{r}
install.packages('textclean')
install.packages('plotly')
install.packages('reshape2')
```


```{r}
library(tidyverse)
library(textclean)
library(lubridate)
library(ggplot2)
library(data.table)
library(zoo)
library(tseries)
library(forecast)
library(scales)
library(plotly)
library(reshape2)
library(tidyr)
library(dplyr)
```


```{r}
df$region <- strip(df$region) 
US_avocados <- df[df$region=="totalus",] ## Isolate rows only containing totalus
avocado_price<- US_avocados[US_avocados$type=="conventional",] ## Isolate only conventional avocados
avocado_price <- avocado_price[order(as.Date(avocado_price$Date, format = "%d/%m/%Y")),]
head(avocado_price)
```


```{r}
#visualising the time series
ggplot()+
  geom_line(data = avocado_price, aes(x = Date, y = AveragePrice),color = "#00AFBB", size = 0.6)+
  geom_vline(xintercept=c(as.Date("2017-09-10"), as.Date("2018-03-25")), linetype="dashed", size =1) +
  annotate("rect", xmin = as.Date("2017-09-10"), xmax = as.Date("2018-03-25"), ymin = -Inf,
           ymax = Inf, alpha =0.1, fill = "red")+
  xlab("Year") + ylab("Avocado Price($)")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.position = "none")
```

```{r}
## Decomposing the timeseries to visualise weeks
week_price <- ts(na.omit(avocado_price$AveragePrice), frequency = 52)
decomp <- stl(week_price, s.window = "periodic")
plot(decomp, cex = 0.5)
```


```{r}
avo_price <- avocado_price$AveragePrice
```

### As p>0.05, the alternative hypothesis cannot be rejected. Therefore, differencing was used to remove trends and seasonal patterns
```{r}
adf.test(avo_price, alternative = "stationary")
```



```{r}
avo_price.diff <- diff(avocado_price$AveragePrice, difference = 1)
adf.test(avo_price.diff,alternative = "stationary")
```


```{r}
auto.arima(avo_price, d=1, seasonal = TRUE)
fit.price <- arima(avo_price[1:140], order = c(0,1,0), seasonal = list(order=c(0,1,1), period = 1))
pre.week <- predict(fit.price, n.ahead = 3)
matplot(1:29, cbind(avo_price[141:169], pre.week$pred), type = "l", xlab = "Weeks", ylab = "Average Price across the US ($)", main = "autoARIMA")
```


```{r}
arima_model_cv <- auto.arima(conv_ts, d=1, D=1, stepwise=FALSE, approximation=FALSE, trace=TRUE)
arima_model_or <- auto.arima(org_ts, d=1, D=1, stepwise=FALSE, approximation=FALSE, trace=TRUE)


print(summary(arima_model_cv))
checkresiduals(arima_model_cv) + theme_minimal()
```


```{r}
```

```{r}
```


```{r}
```